### 2019-01-24 proposal/FINESST/Su_FINESST_2019.tex
FINESST proposal draft v1

    \documentclass[12pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{article}
    \usepackage{
        amsmath,
        amssymb,
        times,
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[margin=1in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

    \newtheorem{theorem}{Theorem}[section]

    \let\Re\undefined
    \let\Im\undefined
    \DeclareMathOperator{\Res}{Res}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \DeclareMathOperator{\diag}{diag\;}

    \DeclarePairedDelimiter\bra{\langle}{\rvert}
    \DeclarePairedDelimiter\ket{\lvert}{\rangle}
    \DeclarePairedDelimiter\abs{\lvert}{\rvert}
    \DeclarePairedDelimiter\ev{\langle}{\rangle}
    \DeclarePairedDelimiter\p{\lparen}{\rparen}
    \DeclarePairedDelimiter\s{\lbrack}{\rbrack}
    \DeclarePairedDelimiter\z{\lbrace}{\rbrace}

    % \everymath{\displaystyle} % biggify limits of inline sums and integrals
    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
    \definecolor{commentgreen}{rgb}{0,0.6,0}
    \lstset{
        basicstyle=\ttfamily\footnotesize,
        frame=single,
        numbers=left,
        showstringspaces=false,
        keywordstyle=\color{blue},
        stringstyle=\color{purple},
        commentstyle=\color{commentgreen},
        morecomment=[l][\color{magenta}]{\#}
    }
    \usepackage[
        backend=bibtex8,
        style=nature]{biblatex}
\defbibenvironment{bibliography}
  {\noindent}
  {\unspace}
  {\printtext[labelnumberwidth]{%
     \printfield{labelprefix}%
     \printfield{labelnumber}}%
   \addspace}
\renewbibmacro*{finentry}{\finentry\addspace}
\bibliography{Su_FINESST_2019}
\AtEveryBibitem{\clearfield{title}}
\begin{document}

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

\singlespacing

\pagestyle{fancy}
\rfoot{Yubo Su}
\rhead{}
\cfoot{\thepage/\pageref{LastPage}}

\title{Nonlinear Tidal Dissipation in Binary White Dwarfs}
\author{Yubo Su}
\date{}

\maketitle

\section{White Dwarf Binaries}

Compact white dwarf (WD) binary systems, with orbital periods in the range of
minutes to hours, are important for several topics in astrophysics. They are
expected to be important sources of gravitational wave (GW) radiation for the
Laser Interferometric Space Antenna (LISA)\cite{lisa}. They are also thought to
be likely progenitors of type Ia supernovae (e.g.\ \cite{Ia0,webbink} or more
recently\cite{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
known, \emph{Gaia} (currently gathering data) is expected to expand the catalog
to a few hundred, and the Large Synoptic Survey Telescope (LSST, first light
scheduled for 2020) is expected to detect a few thousand more\cite{lsst_wd}.
These observations are expected to significantly advance understanding of WD
binaries: results based on \emph{Gaia}'s second data release have already begun
to appear\cite{gaiaDD,gaiaDD2}. The proposed research is well-timed to take
advantage of these new resources.

In spite of the broad importance of WD binaries, the evolution of these systems
pre-merger is sparsely studied. In particular, compact WD binaries are expected
to exhibit major tidal interactions. In these interactions, tidal forces from
the companion WD excite internal gravity waves (IGW), displacements in the WD
fluid experiencing a buoyant restoring force due to density stratification.
As these waves propagate outwards from where they are excited, they grow in
amplitude until they break, as do ocean waves on a shore, and deposit both
energy and angular momentum from the binary orbit in the outer envelope of the
WD\@.

Previous works have shown that this tidal dissipation mechanism can generate
significantly more energy than thermal radiation from the WD surface alone and
is thus a major contributor to the WD energy budget\cite{fullerII,fullerIV}.
However, these works treated the nonlinear wave breaking in a parameterized
fashion. Related works in other fields find that properly including nonlinear
phenomena in hydrodynamical wave breaking produces drastically different IGW
dissipation behavior\cite{winters1994,barker_ogilvie}. Such fully nonlinear
studies have not been performed for WD binaries but are important for the
thermal and orbital evolution of WDs undergoing tidal heating. Characterizing
this nonlinear tidal dissipation will require numerical simulation to capture
the turbulent cascade to small scales that drives IGW dissipation.

\textbf{I propose to study the dynamical effects of tidal dissipation via
nonlinear IGW breaking in binary WDs.} To accomplish this, I aim to address the
following three aims:
\begin{itemize}
    \item Characterize the location, spatial extent, and other properties of
        wave breaking in realistic WD models via direct numerical simulation.
        The location and spatial extent will furnish a simple yet effective
        parameterization of tidal dissipation that can substitute full
        hydrodynamic simulation.

    \item Predict signatures of tidal dissipation over a wide range of possible
        WD systems. In particular, I will describe the impact of tidal heating
        on the apparent temperature of a binary WD and maybe even the production
        of observable flares.

    \item Compute modified GW templates for LISA that account for changes in the
        phase evolution of the orbit due to tidal dissipation.
\end{itemize}

\section{Nonlinear Tidal Dissipation}

\subsection{Background and Preliminary Work}

The current understanding of tidal synchronization in WD binaries is laid out
in~\cite{fullerII}: a tidally-excited train of IGW undergoes wave breaking
in the outer envelope of the WD, locally depositing angular momentum and
synchronizing the WD spin to the orbit. Such a process is broadly consistent
with canonical models of tidal synchronization in stars\cite{zahn75,gn89}, the
only major difference being in the specifics of wave excitation\footnote{While
IGWs in stars are excited at their radiative-convective boundaries, IGWs in WDs
are excited at sharp changes in composition.}. While this tidal synchronization
mechanism is generally accepted, direct numerical simulation of the wave
breaking process is sparse. Since wave breaking is a strongly nonlinear
phenomenon, where a larger wave breaks down into many smaller-scale waves,
numerical simulation is paramount to an accurate understanding of IGW breaking.

IGW breaking is well studied in atmospheric sciences. The wave breaking process
proceeds as follows: initially, the universal instability of an IGW to the
parametric subharmonic instability is responsible for transfering energy and
angular momentum from the unstable wave to the mean flow of the
fluid\cite{drazin}. Subsequently, after the mean flow velocity reaches the
horizontal phase velocity of the IGW, a well-known calculation shows that the
IGW is nearly completely absorbed by the mean flow in the linear
approximation and endowing the atmosphere with a mean horizontal
flow\cite{booker_bretherton,hazel}. However, when this mean flow absorption was
numerically studied including full nonlinear interactions, \emph{new phenomena
not described by the linear theory were observed}\cite{jones_num,winters1994}.
This highlights the importance of numerical simulation in capturing the wave
breaking process.

To this end, I have adapted the spectral hydrodynamic code Dedalus\cite{dedalus}
to simulate a plane-parallel IGW wavetrain excited at the bottom of a 2D
incompressible fluid. A spectral code like Dedalus is ideal for simulating
complex hydrodynamical phenomena as spectral methods have no inherent numerical
viscosity and so are more suitable for resolving the nonlinear cascade to small
length scales in wave breaking.

In my simulations, I observe IGW propagating upwards from where they are
excited, then breaking and depositing horizontal momentum in the fluid, causing
the fluid to acquire average horizontal momentum, consistent with
literature\cite{fullerII}. I derive simple formulae for the location and spatial
extent of the dissipation zone where the IGW is absorbed by the fluid. More
interestingly, I observe partial reflection of the IGW at the synchronization
layer\cite{me}, a phenomenon not considered in the current astrophysical
literature but consistent with atmospheric science results\cite{winters1994}. A
sample simulation is presented in \autoref{fig:nl_fluxes}. I am preparing these
results for publication\cite{me}.
\begin{figure}[!h]
    \centering
    \includegraphics[width=0.7\textwidth]{nl_fluxes.png}
    \caption{The evolution of the horizontal momentum flux $S_{px}(z)$ over time
    in a representative simulation. Note that the location where horizontal
    momentum is deposited by wave breaking, the sharp decrease in $S_{px}$,
    moves to lower $z$ at later times; this reflects spinup of the fluid from
    the top down. The incident flux appears to decrease over time as well, but
    this is attributed to partial reflection of the IGW off the critical
    layer. The vertical axis is in units of the total excited horizontal momentum
    flux, while distances and time are in units of the stratification scale
    height and the buoyancy period.}\label{fig:nl_fluxes}
\end{figure}

\subsection{Proposed Work}

It is therefore clear that full numerical modeling of IGW breaking is necessary
to fully characterize tidal dissipation. The first aim of my proposal is then:
\textbf{I will extend my preliminary results to characterize the dynamics of
nonlinear IGW breaking in realistic WDs.} Via numerical simulation, I will
develop simple models for tidal dissipation that can be used to study WD
evolution under tidal heating without performing further simulations. I will
continue my work in the following stages:
\begin{itemize}
    \item I will perform simulations examining the validity of my results in
        more realistic geometries, including polar and spherical geometries. I
        will continue to use Dedalus, which also supports these coordinate
        systems.

    \item I will extend my simulations to realistic WD models and equations of
        state such as those in~\cite{brassard1992}, continuing to track the
        location and spatial extent of the dissipation layer as well as any new
        phenomena. As WDs vary widely in composition and effective temperature,
        studying representative WD models is vital to obtaining a robust
        characterization of tidal dissipation. My selected WD models are those
        used in the current literature\cite{fullerII, fullerIV}.

    \item With these simulations, I will build a model of tidal dissipation
        valid for a broad range of WDs. I will package my model as a module for
        the 1D stellar evolution code Modules for Experiments in Stellar
        Astrophysics (MESA)\cite{MESA}. MESA is a popular choice for stellar
        evolution simulation due to its computational speed and extensibility.
        It has been used in previous studies to study the evolution of WDs under
        tidal heating\cite{fullerIV,tidal_novae}. My module will provide a
        parameterized and numerically-validated model of tidal dissipation that
        can be used to quickly and easily model tidal interactions in binary
        WDs.
\end{itemize}

\section{Tidal Heating Thermodynamics}

\subsection{Background}

As discussed earlier, compact WD binaries exhibit a range of observed and
hypothesized optical transients. Tidal novae\cite{tidal_novae}, underluminous
supernovae\cite{underlum}, and Ca-rich fast transients\cite{carich} are all
hypothesized to arise in WD binary systems. Given that tidal heating can become
a significant contributor to a WD's total energy budget, a realistic model of
tidal dissipation is important to understanding a binary WD's thermodynamic
evolution during inspiral.

In~\cite{tidal_novae} (hereafter FL), the authors use MESA to study the
production of tidal novae in binary WDs. A simple two-zone parameterization
where the tidal heat is deposited throughout the outer zone was used to model
tidal dissipation. They find that cool WDs in sufficiently compact binaries
(orbital period $\lesssim 15$ minutes) may incur a thermonuclear detonation of
the hydrogen envelope. WD binaries with such short orbital periods have been
observed, e.g.\ SDSS J065133+284423 has a period of 12.75 minutes\cite{12min}.
FL fits the observed properties of SDSS J065133+284423 to their tidal heating
models and finds evidence for tidal heating of the secondary WD in SDSS
J065133+284423. From their fits, FL are able to estimate the temperature of the
WD in the absence of tidal heating. As WDs form around a well known temperature
and cool predictably over time, accounting for tidal heating equates to a large
correction to the age of the WD\@. This illustrates the importance of accurately
modeling tidal interactions when inferring physical properties.

\subsection{Proposed Work}

With the tidal dissipation models I will develop, I will be able to perform a
simulation similar to that in FL but with a more realistic tidal interaction.
The second aim of my proposal is accordingly: \textbf{I will use my tidal
dissipation profiles to simulate a binary WD undergoing tidal heating and make
comparisons to observational data.} The location of the tidal heating is a key
ingredient in determining whether deposited energy can be efficiently transfered
away or whether the WD experiences sudden detonations. As such, an improved
understanding of the properties of tidal dissipation is important to proper
forecasting of binary WDs' thermodynamic evolution. Moreover, while only a few
sufficiently compact WD binaries were available at the time of writing of FL,
\emph{Gaia} data releases 2 and 3 will provide many more compact WD binaries to
examine for evidence of tidal heating.

Using MESA, I will evolve various WD models undergoing tidal heating. From these
simulations, I will extract the increased temperature of the WDs and make
comparisons to observational data, in particular to WDs in new \emph{Gaia} data
releases. I will also identify the occurrence rate and observational properties
of any predicted optical transients such as tidal novae and attempt to identify
them among existing detected events. Predictions of the occurence rates of such
phenomena are vital to guiding future observations. Finally, comparison of
observational data from known WD binaries to simulations could yield new
insights to the behavior of degenerate matter.

\section{Tidal Heating and LISA}

As discussed earlier, WD binaries are a primary source of GW radiation for
LISA\@. LISA will attain optimal sensitivity at frequencies
$10^{-4}$--$10^{-1}\;\mathrm{Hz}$\cite{LISA_band}. Exactly in this frequency
range, tidal effects act to synchronize the orbit of the WDs to the binary
orbit and transfer energy from the orbit into the WDs. While the decay
of the binary orbit is still mostly driven by GW radiation, the tidal energy
dissipation rate grows to $\sim10^{-2}$ the GW energy
dissipation\cite{fullerII,fullerIV}. An effect of such a magnitude causes the
phase of the emitted GWs to significantly from the point-mass binary
prediction; the emitted wave may exhibit ``missing cycles'' due to tidal
effects\cite{fullerII}. GW astronomy uses matched filtering, where a library of
template waveforms is matched against instrument data, to identify GW signals.
As such, the accuracy and completeness of the template library is of utmost
importance.

The final aim of my proposal is then: \textbf{I will use my tidal dissipation
model to compute WD binary GW waveforms including tidal dissipation for use in
the LISA detection pipeline}. This aim is much less computationally expensive
than it appears: LISA-band WD binaries can be well described using the Newtonian
dynamics of two co-orbiting point masses\cite{DWD_pointmass}. The resultant GW
emission can then be accurately computed using the weak gravity quadrupole
approximation (see e.g.~\cite{peters,lsst_wd}). Under these two approximations,
the GW waveform can be computed analytically without resorting to numerical
relativity simulations at all, an enormous computational savings. Thus, I will
compute GW waveforms accounting for the additional phase evolution due to tidal
dissipation. I will publish my corrected waveforms for use by LISA and the GW
community.

\section{Project Timeline}

During the first year of work and first half of the second year, I will complete
my tidal dissipation model. I anticipate the extension of my 2D plane-parallel
work to more complex geometries will be complete within the first half year,
while the extension to realistic WD models will take up to a year. I expect that
these two results together will produce one peer-reviewed publication and one
MESA module.

During the following year, I will use my tidal dissipation model to perform MESA
simulations of tidally heated WDs and compare to observational data. I expect my
MESA simulations and extracting appropriate observables to take about half a
year. I then intend to spend another half year analyzing observational data for
potential signatures of tidal heating or tidal novae. I expect both of these
phases will produce peer-reviewed publications.

Finally, in the last six months I will compute WD binary GW templates for use by
the LISA community. This work will also produce one peer-reviewed publication.

\section{Relevance to NASA Objectives}

This project is extremely relevant to the NASA Astrophysics research program.
My work directly relates to (i) the interactions of particles under the extreme
conditions found in astrophysical situations, (ii) how complex systems create
and shape the structure and composition of the universe on all scales, and (iii)
the development of new techniques that can be applied to future major missions.

My work will improve understanding of possible energetic phenomena that can
occur in compact object binaries. My results concerning angular momentum
transfer via IGW breaking will be applicable to astrophysical systems beyond WD
binaries, for instance angular momentum transfer in stars\cite{l_trans_rev}. I
will devise new ways of analyzing astrophysically interesting systems in
\emph{Gaia} and LSST\@. Finally, my GW templates will be important for LISA GW
detection efforts.

My work has direct relevance to NASA missions in the detectability of
astrophysical transients related to WD binaries. For instance, tidal novae are
theorized to have a similar observational signature to dwarf novae which have
been observed with Chandra and the Hubble Space Telescope, among others.

\printbibliography

\end{document}