### 2019-03-03 proposal/FINESST/Su_FINESST_2019.tex
final proposal

    \documentclass[12pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{article}
    \usepackage{
        amsmath,
        amssymb,
        times,
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[margin=1in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

    \newtheorem{theorem}{Theorem}[section]

    \let\Re\undefined
    \let\Im\undefined
    \DeclareMathOperator{\Res}{Res}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \DeclareMathOperator{\diag}{diag\;}

    \DeclarePairedDelimiter\bra{\langle}{\rvert}
    \DeclarePairedDelimiter\ket{\lvert}{\rangle}
    \DeclarePairedDelimiter\abs{\lvert}{\rvert}
    \DeclarePairedDelimiter\ev{\langle}{\rangle}
    \DeclarePairedDelimiter\p{\lparen}{\rparen}
    \DeclarePairedDelimiter\s{\lbrack}{\rbrack}
    \DeclarePairedDelimiter\z{\lbrace}{\rbrace}

    % \everymath{\displaystyle} % biggify limits of inline sums and integrals
    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
    \definecolor{commentgreen}{rgb}{0,0.6,0}
    \lstset{
        basicstyle=\ttfamily\footnotesize,
        frame=single,
        numbers=left,
        showstringspaces=false,
        keywordstyle=\color{blue},
        stringstyle=\color{purple},
        commentstyle=\color{commentgreen},
        morecomment=[l][\color{magenta}]{\#}
    }
    \usepackage[
        backend=bibtex8,
        style=nature]{biblatex}
% \defbibenvironment{bibliography}
%   {\noindent}
%   {\unspace}
%   {\printtext[labelnumberwidth]{%
%      \printfield{labelprefix}%
%      \printfield{labelnumber}}%
%    \addspace}
% \renewbibmacro*{finentry}{\finentry\addspace}
\bibliography{Su_FINESST_2019}
\AtEveryBibitem{\clearfield{title}}
\begin{document}

% \def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{Section}
% \renewcommand*{\appendixautorefname}{\Snospace}
% \renewcommand*{\figureautorefname}{Fig.}
% \renewcommand*{\equationautorefname}{Eq.}
% \renewcommand*{\tableautorefname}{Tab.}

\singlespacing

\pagestyle{fancy}
\rfoot{Yubo Su}
\rhead{}
\cfoot{\thepage/\pageref{LastPage}}

\title{Nonlinear Tidal Dissipation in Binary White Dwarfs}
\author{Yubo Su}
\date{}

\maketitle

\section{Introduction and Goals of Proposed Research}\label{s:1}

\subsection{White Dwarf Binaries}

Compact white dwarf (WD) binary systems, with orbital periods in the range of
minutes to hours, are important for a range of astrophysical problems. They are
the most important sources of gravitational waves (GWs) for the Laser
Interferometric Space Antenna (LISA)\cite{lisa}. They are also thought to
produce interesting optical transients such as underluminous
supernovae\cite{underlum}, Ca-rich fast transients\cite{carich}, and tidal
novae\cite{tidal_novae}. Most importantly, they have been proposed as the likely
progenitors of type Ia supernovae (e.g.~\cite{Ia0,webbink} or more
recently\cite{Ia1,Ia2}). While presently only a few tens of compact WD binaries
are known\cite{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
expand the catalog to a few hundreds\cite{lsst_wd} (results based on
\emph{Gaia}'s second data release have already begun to
appear\cite{gaiaDD,gaiaDD2}), and the Large Synoptic Survey Telescope (LSST,
first light scheduled for 2020) will likely detect a few thousand
more\cite{lsst_wd}. These observations will significantly advance the
understanding of WD binaries and their evolution. My proposed theoretical and
computational research is well-timed to take advantage of these new advances.

In spite of the broad importance of WD binaries, the evolution of these systems
prior to their final mergers is not well understood. Much of this uncertainty
comes from our imprecise understanding of tidal interactions, which play an
important role during a compact WD binary's inspiral\cite{fullerII}. Previous
studies have shown that these interactions manifest as tidal excitation of
internal gravity waves (IGW), waves in the WD fluid restored by the buoyancy
force due to density stratification\cite{fullerI}. As these waves propagate
outwards towards the WD surface, they grow in amplitude until they break, as do
ocean waves on a shore, and transfer both energy and angular momentum from the
binary orbit to the outer envelope of the WD\cite{fullerI,fullerII}.

Previous works have found that the dissipation of IGW can generate significantly
more energy than thermal radiation from the isolated WD surface and is thus a
major contributor to the WD energy budget\cite{fullerII,fullerIV}. However,
these works parameterized the wave breaking process in an ad hoc manner. The
details of dissipation, namely the location and spatial extent of the wave
breaking, affect the observable outcome: dissipation near the surface of the WD
can be efficiently radiated away and brightens the WD, while dissipation
deep in the WD envelope causes an energy buildup that results in energetic
flares\cite{tidal_novae}. Works in other fields based on numerical simulations
show that strongly nonlinear wave breaking behaves differently than predictions
based on linear and weakly nonlinear theory\cite{winters1994,barker_ogilvie}.
Such fully nonlinear numerical simulations have not been performed for WDs.

\subsection{Goals of Proposed Research}

Characterizing the location and spatial extent of tidal dissipation in WD
binaries will require numerical simulation to capture the turbulent cascade to
small scales that causes wave breaking. \textbf{I propose to study the dynamical
effects of tidal dissipation via nonlinear internal gravity wave (IGW) breaking
in binary WDs.} There are three specific objectives of my proposed research:
\begin{itemize}
    \item Objective A\@: Characterize the location, spatial extent, and other
        properties of wave breaking in realistic WD models via direct numerical
        simulation. The location and spatial extent of wave breaking will
        furnish a simple yet effective parameterization of tidal dissipation in
        a range of WD models. \autoref{s:2} describes the steps required to
        acheive this goal.

    \item Objective B\@: Predict signatures of tidal dissipation over a wide
        range of possible WD systems. In particular, I will study the impact of
        tidal heating on the luminosity of WDs in binaries and explore the
        possibility of producing observable flares. \autoref{s:3} details my
        plan to predict observational manifestations of tidal heating.

    \item Objective C\@: Compute modified GW templates for LISA that account for
        changes in the phase evolution of the orbit due to tidal dissipation.
        \autoref{s:4} elaborates on how I will perform this computation.
\end{itemize}

\section{Objective A\@: Nonlinear Tidal Dissipation}\label{s:2}

\subsection{Background and Preliminary Work}

The current understanding of tidal synchronization in WD binaries is laid out
in~\cite{fullerII}: Tidal forces from the companion excites IGWs in the deep
envelope of the WDs. These IGW propagate outwards and undergo wave breaking in
the outer envelope of the WD, locally depositing angular momentum and
synchronizing the WD spin to the binary orbit. A similar process also operates
in binaries consisting of early type stars\cite{zahn75,gn89}, the only major
difference being in the specifics of wave excitation\footnote{While IGWs in
massive stars are excited at their radiative-convective boundaries, the
excitation of IGWs in WDs is more gradual and is associated with sharp
composition changes in the stellar envelope\cite{fullerII}.}. Nevertheless,
direct numerical simulation of the wave breaking process has not been performed
in either of these systems. Since wave breaking is a strongly nonlinear
phenomenon, where a larger wave breaks down into many smaller-scale waves,
numerical simulation is paramount to an accurate understanding of the tidal
dissipation process.

IGW breaking has been studied in atmospheric sciences. The wave breaking process
proceeds as follows: Initially, as the IGW reaches nonlinear amplitudes, it
breaks down via the parametric instability and transfers energy and angular
momentum from the wave to the mean flow of the fluid\cite{drazin}. After the
mean flow velocity reaches the horizontal phase velocity of the IGW, a critical
layer forms. Analytical calculations show that the IGW is nearly completely
absorbed at this critical layer in the linear approximation and endows the
atmosphere with a mean horizontal flow\cite{booker_bretherton,hazel}. However,
when this mean flow absorption was numerically studied including full nonlinear
interactions, new phenomena not described by the linear theory (reflection off
the critical layer and sharpening of the mean flow) were
observed\cite{jones_num,winters1994}; this nonlinear behavior significantly
affected the evolution of the atmosphere over time. \emph{This highlights the
importance of numerical simulation in capturing the wave breaking process.}

To gain insight into the tidal dissipation process, I have already begun to
adapt the spectral hydrodynamics code Dedalus\cite{dedalus} to study IGW
breaking in a 2D isothermal, stratified atmosphere. A spectral code like Dedalus
is ideal for simulating complex hydrodynamical phenomena, as spectral methods
have no inherent numerical viscosity and so can more accurately resolve the
nonlinear cascade to small length scales in wave breaking.

Working with Dr.\ Daniel Lecoanet (a post-doc at Princeton University, and one
of the authors of the Dedalus code) and my advisor, Prof.\ Dong Lai, I have
simulated the nonlinear evolution of an upward-propagating IGW wavetrain excited
at the bottom of the atmosphere (see \autoref{fig:nl_fluxes} for an example). My
simulations show that waves break and deposit horizontal momentum in the fluid,
causing the fluid to acquire an average horizontal flow, consistent with
previous studies\cite{fullerII}. I have derived simple formulae for the location
and spatial extent of the dissipation zone where the IGW is absorbed by the
fluid. More interestingly, my simulation reveals a partial reflection of the IGW
at the critical layer\cite{me}, a phenomenon not considered in the current
astrophysical literature but consistent with the aforementioned
results\cite{winters1994}. I am preparing these results for
publication\cite{me}.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{nl_fluxes.png}
    \caption{A sample numerical simulation of the evolution of the nonlinear
    internal gravity wave (IGW) propagating upwards and breaking in a stratified
    isothermal atmosphere. The upper panel shows the evolution of the average
    horizontal flow velocity of the fluid $U_0$ (in units of $c_{ph,x}$ the
    horizontal phase velocity of the IGW), and the lower panel shows the
    horizontal momentum flux contained in the IGW $S_{px}(z)$ (in units of the
    total excited flux $S_0$). Different lines correspond to different times in
    the simulation. Times are measured in units of $N^{-1}$, the inverse of the
    Brunt-V\"ais\"al\"a frequency (or the buoyancy period), and heights in units
    of $H$ (the scale height of the density stratification). IGWs are excited at
    $z = 2H$ and propagate to higher $z$ before undergoing wave breaking. The
    sharp decrease in $S_{px}$ corresponds to the location of the critical layer
    where the IGW breaks and drives $U_0$ towards $c_{ph, x}$; it moves to lower
    $z$ over time. The apparent decrease in $S_{px}$ at later times indicates
    wave reflection off the critical layer.}\label{fig:nl_fluxes}
\end{figure}

\subsection{Proposed Work}

It is clear that full numerical modeling of IGW breaking is necessary to fully
capture tidal dissipation. The first aim of my proposal is to \textbf{extend my
preliminary results concerning IGW breaking in stratified atmospheres to
characterize the dynamics of nonlinear IGW breaking in realistic WD models.} Via
numerical simulation, I will develop models for tidal dissipation that can be
used to study long-term WD evolution under tidal heating. I will continue my
work in the following stages:
\begin{itemize}
    \item I will perform simulations examining the validity of my results in
        spherical geometries to capture tidal effects in WD binaries. I will
        continue to use Dedalus, which supports spherical coordinates. Although
        3D simulations are necessarily more complex than my 2D simulations of
        stratified atmospheres (see above), the underlying dynamics of IGW
        breaking are the same.

    \item I will extend my simulations to realistic WD models and equations of
        state such as those in~\cite{brassard1992} as well as those generated by
        MESA\cite{MESA} (see \autoref{s:3} below), continuing to track the
        location and spatial extent of the dissipation layer as well as any new
        phenomena. As WDs vary widely in composition and effective temperature,
        studying representative WD models is vital to obtaining a robust
        characterization of tidal dissipation.
\end{itemize}

\subsection{Feasibility and Availability of Resources}

In my preliminary 2D simulations, I use $256$ horizontal and $1024$ vertical
spectral modes. These simulations take about $2$ days to run on $32$ threads using
Dedalus\cite{dedalus} using adaptive timestepping. Based on these preliminary
results, I intend to perform 3D simulations at resolution $128 \times 128 \times
512$. Including some optimizations I am still developing, I anticipate these
simulations will take about 2 weeks each when performed using 64 threads.

The computational resources available to me at Cornell include a cluster with
104 cores/208 threads as well as multiple 32 core/64 thread machines. There are
sufficient to perform many such 3D simulations within the timespan of the
proposal.

\section{Objective B\@: Tidal Heating and Binary White Dwarf Evolution
}\label{s:3}

\subsection{Background}

As discussed earlier, compact WD binaries may exhibit a range of transient
phenomena: tidal novae\cite{tidal_novae}, underluminous
supernovae\cite{underlum}, and Ca-rich fast transients\cite{carich} are all
hypothesized to arise in WD binary systems. Given that tidal heating can become
a significant contributor to a WD's total energy budget, a realistic model of
tidal dissipation is important to understanding the thermal evolution of WDs
during their binary inspiral.

In~\cite{tidal_novae} (hereafter FL), the authors used MESA\cite{MESA} to study
the production of tidal novae in binary WDs. A simple two-zone parameterization
where the tidal heat is deposited throughout the outer zone was used to model
tidal dissipation. It was found that cool WDs in sufficiently compact binaries
(orbital period $\lesssim 15$ minutes) may incur a thermonuclear detonation of
the hydrogen envelope. WD binaries with such short orbital periods have been
observed, e.g.\ SDSS J0651 has a period of 12.75 minutes\cite{12min}. FL fitted
the observed properties of J0651 to their tidal heating models and finds
evidence for tidal heating of the secondary WD in this system.

\subsection{Proposed Work}

With the tidal dissipation models I will develop in Objective A (see
\autoref{s:2}), I will be able to perform a binary WD evolution study similar to
that in FL but with a realistic tidal dissipation profile (instead of their
parameterized model). The second aim of my proposal is to \textbf{use my tidal
dissipation profiles to simulate a binary WD undergoing tidal heating and make
comparisons to observational data.} The location of the tidal heating is a key
ingredient in determining whether the deposited tidal energy can be efficiently
transfered away or whether the WD experiences sudden detonations. As such, an
improved understanding of the properties of tidal dissipation is important to
proper forecasting of binary WDs' thermodynamic evolution. Moreover, while only
a few sufficiently compact WD binaries were available at the time of writing of
FL, \emph{Gaia} data releases 2 and 3 will provide many more compact WD binaries
to examine for observational manifestations of tidal heating.

Using MESA, I will evolve various WD models undergoing tidal heating. From these
studies, I will extract the increased temperature of the WDs and make
comparisons to observational data, in particular to WDs in new \emph{Gaia} data
releases. I will also identify the occurrence rate and observational properties
of any predicted optical transients such as tidal novae and attempt to identify
them among existing detected events. Predictions of the occurence rates of such
phenomena are vital to guiding future observations. Finally, comparison of
observational data from known WD binaries to my theoretical predictions could
yield new constraints on the interior properties of WDs (e.g.\ crystallization).

As MESA is a 1D stellar evolution code, I will easily have sufficient
computational resources to carry out this investigation.

\section{Objective C\@: Tidal Dissipation and LISA}\label{s:4}

As discussed in \autoref{s:1}, WD binaries are an important source of GW
radiation for LISA\@. LISA will attain optimal sensitivity at frequencies
$10^{-4}$--$10^{-1}\;\mathrm{Hz}$\cite{LISA_band}. Exactly in this frequency
range, tidal effects act to synchronize the spin of the WDs to the binary orbit
and transfer energy from the orbit into the WDs. While the decay of the binary
orbit is still mostly driven by GW radiation, the tidal energy dissipation rate
grows to $\sim10^{-2}$ the GW luminosity\cite{fullerII,fullerIV}. An effect of
such a magnitude causes the phase of the emitted GWs to deviate significantly
from the point-mass binary prediction; the emitted wave may exhibit ``missing
cycles'' due to tidal effects\cite{fullerII}. GW astronomy uses matched
filtering, where a library of template waveforms is matched against instrument
data, to identify GW signals. As such, the accuracy and completeness of the
template library is of utmost importance.

The final aim of my proposal is \textbf{to use my tidal dissipation model to
compute WD binary GW waveforms including tidal dissipation for use in the LISA
detection pipeline}. This aim is much less computationally expensive than it
appears: LISA-band WD binaries can be well described using leading-order
post-Newtonian dynamics and including the effect of tidal
deformations.\cite{DWD_pointmass}. The resultant GW emission can then be
accurately computed using the weak gravity quadrupole approximation (see
e.g.~\cite{peters,lsst_wd}). Under these two approximations, the GW waveform can
be computed analytically without resorting to numerical relativity simulations
at all. Thus, I will compute GW waveforms accounting for the additional phase
evolution due to tidal dissipation. I will publish my corrected waveforms for
use by LISA and the GW community.

\section{Project Timeline}

During the first year of work and first half of the second year, I will complete
calculations of tidal dissipation models. I anticipate that the extension of my
2D plane-parallel work to 3D spherical geometries will be complete within the
first half year, while the extension to realistic WD models will take up to a
year. I expect that these two results together will produce two peer-reviewed
publications in addition to the one currently in preparation.

During the following year, I will use my tidal dissipation model to perform MESA
simulations of tidally heated WDs and compare to observational data. I expect my
MESA simulations and extracting appropriate observables to take about half a
year. I then intend to spend another half year analyzing observational data for
potential signatures of tidal heating or tidal novae. I expect both of these
phases will produce peer-reviewed publications.

Finally, in the last six months I will compute WD binary GW templates for use by
the LISA community. This work will also produce one peer-reviewed publication.

\section{Relevance to NASA Objectives}

This project is extremely relevant to the NASA Astrophysics research program.
My work directly relates to (i) the interactions of particles under the extreme
conditions found in astrophysical situations, (ii) how complex systems create
and shape the structure and composition of the universe on all scales, and (iii)
the development of new techniques that can be applied to future major missions.

My work will improve understanding of possible energetic phenomena that can
occur in compact object binaries. My results concerning angular momentum
transfer via IGW breaking will be applicable to astrophysical systems beyond WD
binaries, for instance angular momentum transfer in massive
stars\cite{l_trans_rev}. I will devise new ways of analyzing astrophysically
interesting systems in \emph{Gaia} and LSST\@. Finally, my GW templates will be
important for LISA GW detection efforts.

My work has direct relevance to NASA missions in the detectability of
astrophysical transients related to WD binaries. For instance, tidal novae are
theorized to have a similar observational signature to dwarf novae which have
been observed with Chandra and the Hubble Space Telescope, among others.

\clearpage

\printbibliography

\end{document}
% mesa school?
% finish FINESST
% e disp? plot

% Kushnir?
% Bouss?? maybe